---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true  
---  
## Loading and preprocessing the data   
```{r load_and_preprocess, echo = TRUE}
library(dplyr)
library(ggplot2)
library(data.table)
library(lubridate)
	
	## Read activity.csv file
	activity_data <- read.csv("./activity.csv", header = TRUE,colClasses = c("numeric", "character", "numeric"))

  ## Convert interval to time format
  activity2 <- cbind(activity_data,                
    as.POSIXct(strptime(paste((activity_data$interval %/% 60) %% 24, ":",activity_data$interval %% 60, sep=""), "%H:%M")))
  
  ## Assign a new name for the time interval column
  colnames(activity2) <- c("steps", "date", "interval", "time_int")

	## Group data by date
	group_day<- group_by(activity2, date)  
```
## What is mean total number of steps taken per day?  
### Total per day   
``` {r print_total}
  ## Summarize data by day
  sum_day <- summarize(group_day, total_steps = sum(steps, na.rm = TRUE))

  ## Create a histogram with total number of steps taken each day
  hist(sum_day$total_steps, xlab = "Total Steps", main = "Histogram of Total Steps Taken Each Day") 
```  
---  
### Total Mean and Median per day
``` {r print_table, results="asis"}
  library(xtable)
  ## Get mean and median
  mm_day <-summarize(group_day, 
    				mean = mean(steps, na.rm=TRUE),
						median = median(steps, na.rm=TRUE)) 
  xt <- xtable(mm_day)
  ## Print table with mean and median
  print(xt, type = "html")
```
### Total Mean per day   
```{r print_mean}

	## Create a graph for Mean Total Steps
	e <- ggplot(data=mm_day, aes(x=date, y=mean))
	
	e + geom_bar(stat="identity", color= "blue", fill="yellow") + 
	  theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5)) 
```  
---  
### Total Median per day  
```{r print_median}
  ## Create a graph for Median Total Steps
  f <- ggplot(data=mm_day, aes(x=date, y=median))
	
	f + geom_bar(stat="identity", color= "blue", fill="yellow") + 
	  theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5)) 
```

## What is the average daily activity pattern?
### Average by interval  
```{r print_ave_by_int}
  ## Group data by time interval
  group_interval <-group_by(activity2, time_int)
	ave_interval <- summarize(group_interval, ave_steps = mean(steps, na.rm = TRUE))

  ## Group data by 5-min interval
  group5_interval <-group_by(activity2, interval)
  ave5_interval <- summarize(group5_interval, ave_steps = mean(steps, na.rm = TRUE))

	
	### Create a time series plot for average number of steps taken
	plot(ave_interval$time_int, ave_interval$ave_steps, xlab = "Interval", ylab = "Average Steps Taken", type = "l")
```  

### Get Interval with Maximum Average of steps  
### The highest time interval with the highest nuber of steps across all days was 835.  
```{r print_max_ave_steps}
  ### Get maximum average
  max_interval <- summarize(ave5_interval, max_interval = max(ave_steps, na.rm = TRUE))
	
	### Get the interval with the maximum average and print it
  print("5-min Interval with highest aveerage maximum of steps")
  ave5_interval[ave5_interval$ave_steps == as.numeric(max_interval),]$interval  
```  
## Imputing missing values   
### Count number of rows with missing values   
```{r count_missing_rows}
  ### Count rows with date == NA
  nrow(activity_data[is.na(activity_data$date),])
	
	### Count rows with interval == NA
	nrow(activity_data[is.na(activity_data$interval),])
	
	### Count rows with steps == NA
	nrow(activity_data[is.na(activity_data$steps),])  
```  
### There were missing values in steps for 2304 rows.  
### Fill in missing values with zeroes   
```{r fill_in_missing}
  
  ## Make another copy of the data
  activity_data2 <- activity2
  
  ### Replace missing values with zeroes
  activity_data2[is.na(activity_data2$steps),]$steps<- 0

  ### Count rows with steps == NA
  nrow(activity_data2[is.na(activity_data2$steps),])   
```  

#### Total per day with zeroes for missing values   
``` {r print_total2}
  ## Group data by date
  group_day2<- group_by(activity_data2, date) 

  ## Summarize data by day
  sum_day2 <- summarize(group_day2, total_steps = sum(steps, na.rm = TRUE))

  ## Create a histogram with total number of steps taken each day
  hist(sum_day2$total_steps, xlab = "Total Steps", main = "Histogram of Total Steps Taken Each Day with Zeroes on NAs") 
```  
---  
#### Total Mean and Median per day  with zeroes for missing values
``` {r print_table2, results="asis"}
  ## Get mean and median
  mm_day2 <-summarize(group_day2, 
    				mean = mean(steps),
						median = median(steps)) 
  mm_day3<- cbind(mm_day, mm_day2)
  colnames(mm_day3) <- c("date", "mean", "median", "date", "mean with NA = zeroes", "median with NA = zeroes")
  xt2 <- xtable(mm_day3)
  ## Print table with mean and median
  print(xt2, type = "html")
```
#### Total Mean per day with zeroes for missing values  
```{r print_mean2}

	## Create a graph for Mean Total Steps
	e2 <- ggplot(data=mm_day2, aes(x=date, y=mean))
	
	e2 + geom_bar(stat="identity", color= "blue", fill="yellow") + 
	  theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5)) 
```  
---  
#### Total Median per day with zeroes for missing values  
```{r print_median2}
  ## Create a graph for Median Total Steps
  f2 <- ggplot(data=mm_day2, aes(x=date, y=median))
	
	f2 + geom_bar(stat="identity", color= "blue", fill="yellow") + 
	  theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5)) 
```  
---
#### Comparison between missing values and filled in with zero values  
#####   Data with NAs has blue and yellow  
#####   Data with NAs replaced with zero has yellow and green  
```{r plot_comparison}
  par(mfrow = c(2, 2))
  
  e + geom_bar(stat="identity", color= "blue", fill="yellow") + 
    theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5)) 
  
  e2 + geom_bar(stat="identity", color= "pink", fill="green") + 
    theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5)) 

  f + geom_bar(stat="identity", color= "blue", fill="green") + 
    theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5)) 
  
  f2 + geom_bar(stat="identity", color= "pink", fill="green") + 
    theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5)) 
  
```  
---  
#### There were no difference between the results with missing data and the results with zeroes on missing values.  
## Are there differences in activity patterns between weekdays and weekends?
```{r print_weekday_end}
  library(lubridate)
  library(chron)
	
  ## Copy original data and replace nulls with zeroes
	ad2 <- as.data.table(activity2)
  ad2[is.na(ad2$steps),]$steps <- 0

  ## Add date column
	ad2[,date2:=as.Date(ad2$date, "%Y-%m-%d")]

  ## Add day_type column and initialize it with "weekday"
  ad2 <-cbind(ad2, c("weekday"))
  
  ## Change column names
  setnames(ad2, c("steps","date","interval", "time_int","date2","V2"), c( "steps","date","interval", "time_int","date2","day_type"))

  ## Put "weekend" on day_type for weekend days
  ad2[is.weekend(ad2$date2),]$day_type="weekend"

  ## Group by Weekend indicator and interval and then calculate mean	
	group_interval2 <-group_by(ad2,  day_type, interval)
	ave_interval2 <- summarize(group_interval2, ave_steps = mean(steps))
	
	## Create a time series plot for average number of steps taken with day_type
 qplot(interval, ave_steps, data = ave_interval2, facets = day_type ~., xlab = "Interval", ylab = "Aveage Steps", geom = (c("point", "line")))

```